\documentclass{ximera}
\title{Initial Assessment}
\author{Oguz Kurt}

\begin{document}
\begin{abstract}
This is an old final exam for this course. Please, answer the following questions seriously so that I will have an understanding of your current level. Make sure to not look into any online or paperback references while you are trying these problems. It is essential for me to have an understanding about your current level. I will base my teaching on this. Please, do not use calculator unless you cannot do a very basic calculation. 
\end{abstract}
\maketitle

\begin{problem}
Perform the indicated operation and simplify: \hspace{1cm} 
$$ 7(-6) - 4 \cdot (2 - 5) =\answer{-30}$$


\end{problem}



\begin{problem}
Distribute and simplify: \hspace{1cm} 
$$3 \left [ 2 (y - 3) - 2 \right ] - 2 (3y + 2)=\answer{-28}$$





\end{problem}



\begin{problem}
Subtract: \hspace{1cm} 
$$\displaystyle \frac{3}{5} - \frac{2}{7}=\answer{\frac{11}{35}}$$





\end{problem}



\begin{problem}
Solve: \hspace{1cm} $4t + 5 < 8 t -3$\\

Give the answer by graphing on a number line and by using interval notation. \\

Interval Notation = $\answer{(2,\infty)}$





\end{problem}



\begin{problem}
Solve: \hspace{1cm} $3(z-5) = 2z-9$\\

$z=\answer{6}$



\end{problem}



\begin{problem}
Solve: \hspace{1 cm}$ -2(3x-5) = 3x - 2(1 + 2x) + 2$ \\

$x=\answer{2}$

\end{problem}



\begin{problem}
Linda spent 3/4 of her savings on furniture and the rest on a TV. If the TV cost her \$200, what were her original savings? 

Her original savings are \$ $\answer{800}$.



\end{problem}



\begin{problem}
Find the slope of the line that passes through the points $(1,-3)$ and $(-3,5)$.

Slope $m=\answer{-2}$.




\end{problem}



\begin{problem}
Determine the slope and the y-intercept for the line given by $2x - 4y = 20$. Graph the line on a plane.

Slope is $m=\answer{\sage{1/2}}$ and y-intercept is $(\answer{0},\answer{5})$.


\end{problem}



\begin{problem}
Perform the indicated operation and simplify.
$$(3x^5 - 2x^3 +7x^2+5x -11) - (x^5 - 6x^4 + 2x^2 -7)
=
\answer{2x^5+6x^4-2x^3+5x^2+5x-4}
$$




\end{problem}



\begin{problem}

Multiply: \hspace{1 cm} 
$$(2x -5)(3x^2 +2x -6)
=
\answer{6x^3-11x^2-22x+30}
$$



\end{problem}



\begin{problem}

Simplify $\dfrac{4\cdot 2^{200}}{8^{67}}=\answer{2}$.

Show your work. Your answer must be an integer. \\ 



\end{problem}



\begin{problem}

Factor usigin grouping:  
$$x^5 - 2 x^3 - 3x^2 +6
=
(\answer{x^2-2})\cdot(\answer{x^3-3})$$

The first factor is a degree two polynomial and the second factor is a degree three polynomial.


\end{problem}



\begin{problem}

Factor: 
$$49y^2 - 25
=
(\answer{7y-5})\cdot (\answer{7y+5})$$

\end{problem}



\begin{problem}

Solve: \hspace{1cm} $x^2 - 6x = 7$

$$x=\answer{-1},~~~ \answer{7}.$$


\end{problem}



\begin{problem}

Solve: \hspace{1cm} $\dfrac{x^2+x-2}{x+2}= 0$.


$$x=\answer{1}.$$
\end{problem}



\begin{problem}
Solve: \hspace{1cm} $\dfrac{x}{2}+\dfrac{4}{x}= x+1$.

$$x=\answer{-4}, ~~~ \answer{2}.$$

\end{problem}

\end{document}
