\documentclass{ximera}
\input{../preamble.tex}
\title{Discrete Random Variables and Probability Distributions}
\author{Oguz Kurt}

\begin{document}
\begin{abstract}
This homework covers examples of discrete random variables, their distribution as well as expected values, mean and variance of random variables.
\end{abstract}
\maketitle

\begin{sagesilent}
def c(n,k):
    if 0<=k and k <=n:
        return binomial(n,k)
    else:
        return "error"

def p(n,k):
    if 0<=k and k <=n:
        return binomial(n,k)*factorial(k)
    else:
        return "error"
def m(n,mylist):
    l=len(mylist)
    return factorial(n)/(prod([factorial(mylist[i]) for i in range(l)]))
\end{sagesilent}

\section*{Syntax}
Here is a reminder of the way to input some numbers. Please, use the parsing power of your browser instead of using a calculator.

\hspace{0.5cm}


\begin{tabular}{c|c}
Operation & Syntax  
\\
\hline
$a+b$ & {\color{red} \verb!a+b!} 
\\
\hline
$a- b$ & {\color{red} \verb!a-b!} 
\\
\hline
$a\cdot b$ & {\color{red} \verb!a*b! } 
\\
\hline
$a/b=\frac{a}{b}$ & {\color{red} \verb!a/b!}
\\
\hline
$a^b$ & {\color{red} \verb|a**b| or \verb|a^b|} 
\\
\hline
$\sqrt{a}$ & {\color{red} \verb|sqrt(a)| or \verb|a**(1/2)|} 
\\
\hline
$a!$ & {\color{red} \verb|a!|} 
\\
\hline
$\binom{a}{b}$ & {\color{red} \verb|a!/(b!*(a-b)!)|}
\\
\hline
$P_k^n$ & {\color{red} \verb|a!/(a-b)!|}
\\
\hline
$e^x$ & {\color{red} \verb|exp(x)|}
\\
\end{tabular}

\hspace{1cm}

\begin{problem}
Sixty percent of new drivers have had driver education. During their first year, new drivers without driver education have probability .08 of having an accident, but new drivers with driver education have only a .05 probability of an accident. What is the probability a new driver has had driver education, given that the driver has had no accidents the first year? 

\begin{prompt}%err
$$\Pr(\text{had education }|\text{ no accidents})=\answer{\frac{(.95)(.6)}{(.95)(.6)+(.92)(.4)}}$$
\end{prompt}
\end{problem}


\begin{problem}
An urn contains 6 red balls and 3 blue balls. One ball is selected at random and is replaced by a ball of the other color. A second ball is then chosen. What is the conditional probability that the first ball selected is red, given that the second ball was red? 

\begin{prompt}
$$\Pr(\text{1st ball RED }|\text{ 2nd ball RED})=\answer{\frac{10}{17}}$$

\end{prompt}
\end{problem}



\begin{problem}
 A test for a disease correctly diagnoses a diseased person as having the disease with probability .85. The test incorrectly diagnoses someone without the disease as having the disease with probability .10. If 1\% of the people in a population have the disease, what is the chance that a person from this population who tests positive for the disease actually has the disease? 

\begin{prompt}
$$\Pr(\text{has disease }|\text{ tests positive})=\answer{ \frac{(0.85)(0.01)}{(0.85)(0.01)+(0.1)(0.99)}}$$
\end{prompt}
\end{problem}


\begin{problem}
Let $X$ be a random variable with distribution function $p_X(x)=\Pr(X=x)$ defined by
$$p_X(−1) = 1/5,~ p_X(0) = 1/5,~ p_X(1) = 2/5,~ p_X(2) = 1/5.$$

\medskip

\begin{enumerate}
    \item Let $Y$ be the random variable defined by the equation $Y = X + 3$. Find
the distribution function $p_Y (y)$ of $Y$.

\medskip

\begin{prompt}
    Please, write the values of $Y=y$ and $p_Y(y)$ in the following table in an increasing order.
    $$
    \begin{array}{c|c}
    Y=y & p_Y(y) \\
    \hline
    \answer{2} & \answer{\frac{1}{5}} \\
    \hline
    \answer{3} & \answer{\frac{1}{5}} \\
    \hline
    \answer{4} & \answer{\frac{2}{5}} \\
    \hline
    \answer{5} & \answer{\frac{1}{5}} \\
    \end{array}
    $$
\end{prompt}

    \item Let $Z$ be the random variable defined by the equation $Z = X^2$. Find the
distribution function $p_Z(z)$ of $Z$.

\medskip

\begin{prompt}
    Please, write the values of $Y=y$ and $p_Y(y)$ in the following table in an increasing order.
    $$
    \begin{array}{c|c}
    Z=z & p_Z(z) \\
    \hline
    \answer{0} & \answer{\frac{1}{5}} \\
    \hline
    \answer{1} & \answer{\frac{3}{5}} \\
    \hline
    \answer{4} & \answer{\frac{1}{5}} \\
    \end{array}
    $$

\end{prompt}
\end{enumerate}
\end{problem}

\begin{problem}

If the number of typographical errors per page follows a Poisson distribution with parameter $\lambda=0.2$, what is the probability that the total number of errors in 10 randomly selected pages is 10? 

\begin{prompt} %err
$$\Pr(\text{10 errors in 10 pages})= \answer{\sage{exp(-2)*2^10/(factorial(10))}}$$
\end{prompt}
\end{problem}


\begin{problem}
If $X$ has a uniform distribution on $\{0,~1,~2,\ldots,~99,~100\}$, then what is $\Pr\left[X+\frac{1000}{X}\geq 70\right]$? 

\begin{prompt}
$$\Pr\left[X+\frac{1000}{X}\geq 70\right]=\answer{\frac{72}{101}}$$
\end{prompt}
\end{problem}


\begin{problem}
 Consider independent success-failure trials with success probability $p=0.3$. Let $X$ be the number of trials up to and including the third failure. For $k = 3, 4, 5, \ldots$ , Calculate $\Pr[X=11]$ and $\Pr[X\geq 5]$?

\begin{prompt}%err
\begin{eqnarray*}
\Pr[X=11]&=&\answer{\sage{binomial(10,2)*(0.3)^3*(0.7)^8}} \\
\Pr[X\geq 5]&=&\answer{\sage{1-binomial(2,2)*(0.3)^3-binomial(3,2)*(0.3)^3*(0.7)^1-binomial(4,2)*(0.3)^3*(0.7)^2}}
\end{eqnarray*}
\end{prompt}
\end{problem}


\begin{problem}
Suppose $X$ has a Poisson distribution with a standard deviation of 4. What is the conditional probability that $X$ is exactly 1 given that $X\geq1$? 

\begin{hint}
Variance and mean of a Poisson distribution are equal.
\end{hint}

\begin{prompt}
        $$\Pr[X=1~|~X\geq 1] =\answer{\sage{(exp(-16)*16^1)/(1-(exp(-16)*16^0))}}$$
\end{prompt}
\end{problem}

%%%%%%%%%%%
\begin{problem}
A box is to be constructed so that its height is 10 inches and its base is $X$ inches by $X$ inches. If $X$ has a uniform distribution over $\{2,3,4,5,6,7,8\}$, then what is the expected volume of the box in cubic inches? 
\begin{prompt}
Suppose $V$ is the volume of the box with 10 inches heights and $X$ inches by $X$ inches bottom. Then $V=\answer{10x^2}$ whenever $X=x$.
$$E[V]=\answer{\sage{10*(2^2+3^2+4^2+5^2+6^2+7^2+8^2)/7}}
$$
\end{prompt}

\end{problem}



%%%%%%%%%%% TESTED UPTO HERE
\begin{problem} The number of traffic accidents per week in a small city has a Poisson distribution with mean equal to 3. What is the probability of exactly 2 accidents in 2 weeks? 

\begin{prompt}
Let $X\sim Poi(3)$ represent a single week. If we extend the interval to two weeks, we get a new random variable $Y=Poi(\answer{6})$ and the problem asks us to calculate
$$\Pr[Y=2]=\answer{\sage{(exp(-6)*6^2)/(factorial(2))}}.$$
\end{prompt}

\end{problem}

%%%%%%%%%%%
\begin{problem}
In a certain communications system, there is an average of 1 transmission error per 10 seconds. Let the distribution of transmission errors be Poisson. What is the probability of more than 1 error in a communication half-a-minute in duration? 



\begin{prompt}
If $X$ is the Poisson random variable with an average of 1 error per 10 seconds, then we first need a Poisson random variable $Y$ with an average of $\answer{3}$ errors per 30 seconds. We can then calculate:

$$\Pr[Y>1]=\answer{\sage{1-exp(-3)-3*exp(-3)}}$$
\end{prompt}

\end{problem}

%%%%%%%%%%%
\begin{problem}
Let $X$ be a Poisson random variable with mean $\lambda$.If $\Pr[X=1\mid X\leq1]=0.8$,what is the value of $ \lambda$? 
\begin{prompt}
$$\lambda=\answer{4}$$
\end{prompt}

\end{problem}

%%%%%%%%%%%
\begin{problem}
Coins $K$ and $L$ are weighted so the probabilities of heads are .3 and .1, respectively. Coin $K$ is tossed 5 times, and coin $L$ is tossed 10 times. If all the tosses are independent, what is the probability that coin $K$ will result in heads 3 times and coin $L$ will result in heads 6 times? 

\begin{prompt}
$$\Pr[K=3, L=6]=\answer{\sage{(binomial(5,3)*.3^3*.7^2*binomial(10,6)*.1^6*.9^4)}}$$
\end{prompt}

\end{problem}

\end{document}
%%%%%%%%%%%
\begin{problem}
If $Var(X + Y) = 3$, $Var(X-Y)=1$, $E(X) = 1$, and $E(Y) = 2$, then what is $E(XY)$? 
\begin{hint}
\item You might want to calculate $Var(X+Y)-Var(X-Y)$. 
\item $E[A+B]=E[A]+E[B]$ for random variables A and B.
\item $Var(C)=E[C^2]-(E[C])^2$ for random variable C.
\item Using the above two facts above will lead to cancellation of bunch of terms and will leave only the ones you need!
\end{hint}

\begin{prompt}
$$
E[XY]=\answer{10/4}
$$ 
\end{prompt}

\end{problem}

%%%%%%%%%%%
\begin{problem}
Y is a Bernoulli random variable with the probability of heads, $0.4$ and a biased die has the following probabilities for each outcome where $X$ is the number of dots on the outcome. X and Y are independently rolled/flipped. Calculate the following information.
$$\begin{array}{l|c}
X=x & \Pr(X=x) \\
\hline
1& 0.3 \\
2& 0.1\\
3& 0.15\\
4& 0.2\\
5& 0.15\\
6& 0.1 \\
\end{array}$$

\begin{prompt}
\begin{enumerate}
    \item $E[X]=\answer{\sage{0.3+2*0.1+3*0.15+4*0.2+5*0.15+6*0.1}}$
    \item $E[X^2]=\answer{\sage{0.3+2^2*0.1+3^2*0.15+4^2*0.2+5^2*0.15+6^2*0.1}}$
    \item $Var(X)=\answer{\sage{0.3+2^2*0.1+3^2*0.15+4^2*0.2+5^2*0.15+6^2*0.1-(0.3+2*0.1+3*0.15+4*0.2+5*0.15+6*0.1)^2}}$
    \item $E[3X^2+5X]=\answer{\sage{3*(0.3+2^2*0.1+3^2*0.15+4^2*0.2+5^2*0.15+6^2*0.1)+5*(0.3+2*0.1+3*0.15+4*0.2+5*0.15+6*0.1)}}$
    \item $E[1/X]=\answer{\sage{0.3+1/2*0.1+1/3*0.15+1/4*0.2+1/5*0.15+1/6*0.1}}$
    \item $E[X-1/X]=\answer{\sage{(0.3+2*0.1+3*0.15+4*0.2+5*0.15+6*0.1)-(0.3+1/2*0.1+1/3*0.15+1/4*0.2+1/5*0.15+1/6*0.1)}}$
    \item $E[Y]=\answer{0.4}$
    \item $E[X+2Y]=\answer{\sage{0.3+2*0.1+3*0.15+4*0.2+5*0.15+6*0.1+2*0.4}} $
    \item $E[X\cdot Y]=\answer{\sage{1*0*0.3*0.6+1*1*0.3*0.4+2*0*0.1*0.6+2*1*0.1*0.4+3*0*0.15*0.6+3*1*0.15*0.4+4*0*0.2*0.6+4*1*0.2*0.4+5*0*0.15*0.6+5*1*0.15*0.4+6*0*0.1*0.6+6*1*0.1*0.4}}$
    \item $E[Y/X]=\answer{\sage{0/1*0.3*0.6+1/1*0.3*0.4+0/2*0.1*0.6+1/2*0.1*0.4+0/3*0.15*0.6+1/3*0.15*0.4+0/4*0.2*0.6+1/4*0.2*0.4+0/5*0.15*0.6+1/5*0.15*0.4+0/6*0.1*0.6+1/6*0.1*0.4}}$
    \item Covariance of $X$, $Y$: $$Cov(X,Y)=E[X\cdot Y]-E[X]\cdot E[Y]=\answer{0}$$
    %sage{(1*0*0.3*0.6+1*1*0.3*0.4+2*0*0.1*0.6+2*1*0.1*0.4+3*0*0.15*0.6+3*1*0.15*0.4+4*0*0.2*0.6+4*1*0.2*0.4+5*0*0.15*0.6+5*1*0.15*0.4+6*0*0.1*0.6+6*1*0.1*0.4)-(0.3+2*0.1+3*0.15+4*0.2+5*0.15+6*0.1)*(0.4)}}
    \item Correlation coefficient of $X$, $Y$:
    $$r_{X,Y}=\answer{0}$$
    \item Why are $Cov(X,Y$ and $r_{X,Y}$ both equal to $0$?
    \begin{freeResponse}
    
    \end{freeResponse}
\end{enumerate}

\end{prompt}
\end{problem}

%%%%%%%%%%%
\begin{problem}The following table of probabilities are given for the students at the University of Xenia. Suppose that $X=1,2,3,4$ for blood types $A,B,AB,0$, respectively, and $Y=\pm 1$ for Rh$\pm$, respectively, are random variables. Calculate the following:

\begin{tabular}{c|c|c|c|c}
& A & B & AB & 0 \\ 
\hline
Rh- & $0.15$ & $0.22$ & $0.15$ & $0.04$ \\ 
\hline
Rh+ & $0.10$ & $0.10$ & $0.18$ & $0.06$ \\ 
\end{tabular}

\begin{prompt}
\begin{enumerate}
    \item $E[X]=\answer{\sage{1*0.25+2*0.32+3*0.33+4*0.1}}$
    \item $E[X^2]=\answer{\sage{1^2*0.25+2^2*0.32+3^2*0.33+4^2*0.1}}$
    \item $Var(X)=\answer{\sage{(1^2*0.25+2^2*0.32+3^2*0.33+4^2*0.1)-(1*0.25+2*0.32+3*0.33+4*0.1)^2}}$
    \item $E[Y]=\answer{\sage{(+1)*0.44+(-1)*0.56}}$
    \item $E[Y^2]=\answer{\sage{1}}$
    \item $Var(Y)=\answer{\sage{1-((+1)*0.44+(-1)*0.56)^2}}$
    \item $E[X\cdot Y]=\answer{\sage{(-1)*1*0.15+(1)*1*0.10+(-1)*2*0.22+(1)*2*0.10+(-1)*3*0.15+(1)*3*0.18+(-1)*4*0.04+(1)*4*0.06}}$
    \item Covariance of $X$, $Y$: $$Cov(X,Y)=E[X\cdot Y]-E[X]\cdot E[Y]=\answer{\sage{((-1)*1*0.15+(1)*1*0.10+(-1)*2*0.22+(1)*2*0.10+(-1)*3*0.15+(1)*3*0.18+(-1)*4*0.04+(1)*4*0.06)-(1*0.25+2*0.32+3*0.33+4*0.1)*((+1)*0.44+(-1)*0.56)}}$$
    %sage{(1*0*0.3*0.6+1*1*0.3*0.4+2*0*0.1*0.6+2*1*0.1*0.4+3*0*0.15*0.6+3*1*0.15*0.4+4*0*0.2*0.6+4*1*0.2*0.4+5*0*0.15*0.6+5*1*0.15*0.4+6*0*0.1*0.6+6*1*0.1*0.4)-(0.3+2*0.1+3*0.15+4*0.2+5*0.15+6*0.1)*(0.4)}}
    \item Correlation coefficient of $X$, $Y$:
    $$r_{X,Y}=\frac{Cov(X,Y)}{\sigma_{X}\cdot \sigma_{Y}}=\answer{\sage{(((-1)*1*0.15+(1)*1*0.10+(-1)*2*0.22+(1)*2*0.10+(-1)*3*0.15+(1)*3*0.18+(-1)*4*0.04+(1)*4*0.06)-(1*0.25+2*0.32+3*0.33+4*0.1)*((+1)*0.44+(-1)*0.56))/(sqrt((1^2*0.25+2^2*0.32+3^2*0.33+4^2*0.1)-(1*0.25+2*0.32+3*0.33+4*0.1)^2)*sqrt(1-((+1)*0.44+(-1)*0.56)^2))}}$$
    \item Based on $r_{X,Y}$, we can conclude that the random variables $X$ and $Y$ are 
    \begin{multipleChoice}
        \choice{Positively Correlated}
        \choice{Negatively Correlated}
        \choice[correct]{Not Correlated}
    \end{multipleChoice}
    \item Explain your choice in the previous problem?
    \begin{freeResponse}
        
    \end{freeResponse}
\end{enumerate}

\end{prompt}

\end{problem}

%%%%%%%%%%%
\begin{problem}
Let $Y$ and $Z$ be two random variables. If $Var(Y) = 4, Var(Z) = 16 $, and $Cov(Y,Z) = 2$, what is $Var(3Z-2Y)$?

\begin{prompt}
$$
Var(3Z-2Y)=\answer{\sage{136}}
$$
\end{prompt}

\end{problem}

\begin{problem}
An automobile manufacturing company produces three different car models. The table below presents sales data and average gasoline consumption data for these three models. What is the mean miles per gallon (mpg) for the cars sold by the company, assuming each car uses the same number of gallons of gasoline?

\begin{tabular}[hc]{l|c|c}
Model & \# of cars sold & mpg \\
\hline
I & 2000 & 16 \\
\hline
II & 2800 & 20 \\
\hline
III & 3200 & 25 \\
\end{tabular}
\begin{prompt}
$$
\mu=\answer{\sage{(2000*16+2800*20+3200*25)/(2000+2800+3200)}}
$$
\end{prompt}

\end{problem}

\begin{problem}
An experiment is repeated 50 times with 20 successful outcomes. The designers use a random variable $X$ which counts the number of successful experiments from a randomly selected set of 15. What is the probability that they will select at most 2 successful experiments $(X\leq 2)$.  
\begin{prompt}
$$
\Pr(X\leq 2)=\answer{\sage{(binomial(20,0)*binomial(30,15)+binomial(20,1)*binomial(30,14)+binomial(20,2)*binomial(30,13))/(binomial(50,15))}}
$$
\end{prompt}

\end{problem}

\end{document}

