\documentclass{ximera}
\input{../preamble.tex}
\author{Oguz Kurt}

\title{Exam 2 problems}
\begin{document}
%feel free to change the title page
%

\begin{abstract}
\empty
\end{abstract}
\maketitle

\begin{problem}
The first problem is covered directly in HW.

\end{problem}

\begin{problem}

\begin{enumerate}
    \item A grasshopper hops each time with a probability of $0.99$. What is the probability that it will fail to hop at most 3 times in 100 tries?  
\begin{prompt}
$$X\sim Binomial(10,0.99)\Longrightarrow\Pr(X\leq 3)=\answer[tolerance=0.000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000001]{\sage{binomial(100,0)*(0.99)^0*(0.01)^100+binomial(100,1)*(0.99)^1*(0.01)^99+binomial(100,2)*(0.99)^2*(0.01)^98+binomial(100,3)*(0.99)^3*(0.01)^97}}$$
\end{prompt}
\item We know that a honey bee dies after stinging somebody. Suppose probability that a honeybee will sting somebody when felt in danger is $0.1$. What is the probability that it will die in {\bf at most} the 3rd encounter with someone posing danger? 
\begin{prompt}
$$Y\sim Geometric(0.1)\Longrightarrow\Pr(Y\leq 3)=\answer[tolerance=0.000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000001]{0.1+0.1*0.9+0.1*0.9^2}$$
\end{prompt}
\item A barrel has 5 poisonous snakes and 45 non-poisonous snakes. You pick 6 snakes randomly. What is the probability that you will pick at least one poisonous snake? Is it less than 1/2?
\begin{prompt}
$$Z\sim HyperGeo(50,5,6)\Longrightarrow\Pr(Z\geq 1)=1-\Pr(Z=0)=\answer[tolerance=0.000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000001]{\sage{1-(binomial(45,6))/(binomial(50,6))}}$$
\end{prompt}
\end{enumerate}
\end{problem}


%\newpage

\begin{problem}
{\bf (10 points each)} \\
\large Please, do NOT simplify your answers!
\begin{enumerate}
    \item The number of defective engines produced in a {\bf day} by a car manufacturer is a Poisson random variable with parameter 5. What is the probability that the factory will produce at most 3 defective engines in a {\bf 5-day-week}?
\begin{prompt}    
    $X\sim Poi(5)$ on an interval size of $1$ day. We use a $5-day$ interval. So, we need to use $Y\sim Poi(\answer{25})$ as our random variable and solve
    $$\Pr(Y\leq 3)=\answer[tolerance=0.000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000001]{\sage{exp(-25)*(1+25+25^2/2+25^3/6)}}$$
\end{prompt}

    \item The number of non-engine defective car parts produced in a {\bf 5-day-week} by the same manufacturer is a Poisson random variable with parameter 15. What is the probability that the company will produce at least 3 cars with non-engine defects in a single day?
\begin{prompt}    
    $Z\sim Poi(15)$ on an interval size of $5$ days. We use a 
    $1-day$ interval. So, we need to use $W\sim Poi(\answer{3})$ as our random variable and solve
    $$\Pr(W\leq 3)=1-\Pr(W=2)-\Pr(W=1)-\Pr(W=2)=\answer[tolerance=0.000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000001]{\sage{1-exp(-3)*(1+3+3^2/2)}}$$
\end{prompt}    

    \item {\bf (Bonus)} What is the probability that the company will produce exactly 10 defective (engine or non-engine) cars in a {\bf 5-day-week}?
\begin{prompt}    
We count the total number of defects in a 5-day interval. So, we need to use a random variable $D=Y+Z$ where $Y\sim Poi(25)$ and $Z\sim Poi(15)$ on the same 5-day interval size. Hence, $D\sim Poi(\answer{25+15})$. The answer then is
    $$\Pr(Z=10)=\answer[tolerance=0.000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000001]{\sage{exp(-40)*40^10/factorial(10)}}$$
\end{prompt}    


\end{enumerate}
\end{problem}



\begin{problem} { \bf (10 points each)} 
Let X be a random variable counting the number of dots on a fair die after a single roll and Y be the total number of dots after 3 independent rolls of the same die.  Calculate the following:
\\
{\bf Hint:} You might want to first calculate the mean and variance of $X$ in the extra space below. Then use the fact that $Y=X_1+X_2+X_3$ where $X_i$ is the number of dots in the $i^{\text{th}}$ roll of the die.


\begin{explanation}
\begin{itemize}
    \item $Y\sim X_1+X_2+X_3$
    \item We need to calculate the mean and variance of $X$ first.
 
 $E[X]=\answer{21/6}$ and $Var(X)=\answer{((1-3.5)^2+(2-3.5)^2+(3-3.5)^2+(4-3.5)^2+(5-3.5)^2+(6-3.5)^2)/6}$
    \item We know $E[A+B]=E[A]+E[B]$ for all random variables.
    \item We know $Cov(A,B)=0$ whenever A and B are independent.
    \item We know $Cov(A,A)=Var(A)$ for any random variable.
    \item We know
$$Var[X_1+X_2+X_3]=Var[X_1]+Var[X_2]+Var[X_3]+2Cov(X_1,X_2)+2Cov(X_1,X_3)+2Cov(X_2,X_3)   $$
\end{itemize}
\end{explanation} 

\begin{enumerate}
    \item $\mathbb E[Y]=\answer{3*21/6}$
    
    \item $\mathbb Var[Y]=\answer{3*(1-3.5)^2+(2-3.5)^2+(3-3.5)^2+(4-3.5)^2+(5-3.5)^2+(6-3.5)^2)/6}$


    \item {\bf (Bonus)} Is $\mathbb Var[Y]=\mathbb Var[3X]$? Why?
    \begin{multipleChoice}
        \choice{True}
        \choice[correct]{False}
    \end{multipleChoice}
    because $Var(Y)=\answer{3}\cdot Var(X)$ but $Var(3X)=\answer{9}\cdot Var(X)$.
    \vspace{2cm}
\end{enumerate}
\end{problem}
\end{document}
