\documentclass{ximera}
\input{../preamble.tex}
\title{Basic Definitions}
\author{Oguz Kurt}
\begin{abstract}
   Basic definitions needed to start statistics. Copied from OpenStax.org to help with the first HW.
\end{abstract}

%\outcome{Identify, distinguish natural, whole numbers and integers.}
%\outcome{Use natural, whole numbers and integers in operations.}
%\outcome{Learn the impact of different operations on natural, whole numbers and integers.}

\begin{document}
\maketitle

\begin{itemize}
\item \textbf{Variable:}
a characteristic of interest for each person or object in a population
\item \textbf{Categorical Variable:}
variables that take on values that are names or labels. Nominal is non-number and ordinal is a number variable that is categorical.
\item \textbf{Numerical Variable:}
variables that take on values that are indicated by numbers. If the order of the input matters but not the proportion, it is an interval type variable. Otherwise, it is a ratio type variable. 
\item \textbf{Data:}
a set of observations (a set of possible outcomes); most data can be put into two groups: qualitative (an attribute whose value is indicated by a label) or quantitative (an attribute whose value is indicated by a number). Quantitative data can be separated into two subgroups: discrete and continuous. Data is discrete if it is the result of counting (such as the number of students of a given ethnic group in a class or the number of books on a shelf). Data is continuous if it is the result of measuring (such as distance traveled or weight of luggage)
\item \textbf{Population:}
all individuals, objects, or measurements whose properties are being studied
\item \textbf{Sample:}
a subset of the population studied
\item \textbf{Representative Sample:}
a subset of the population that has the same characteristics as the population
\item \textbf{Proportion:}
the number of successes divided by the total number in the sample
\item \textbf{Parameter:}
a number that is used to represent a population characteristic and that generally cannot be determined easily
\item \textbf{Average:}
also called \textbf{mean}; a number that describes the central tendency of the data
\item \textbf{Probability:}
a number between zero and one, inclusive, that gives the likelihood that a specific event will occur
\item \textbf{Statistic:}
a numerical characteristic of the sample; a statistic estimates the corresponding population parameter.
\end{itemize}
\end{document}
